---
title: "HW2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, include =FALSE}
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
tidymodels_prefer()
```

```{r, include = FALSE}
ab <- read_csv("abalone.csv")
```

### Question 1: 

```{r, include =TRUE}
#Creating function ab1 which calculates age using rings + 1.5 and then adding it to the abalone dataset 
ab1 <- ab %>% 
  mutate(age = (ab$rings + 1.5)) %>% 
  select(everything(), age)

#Plot a bar graph to be able to better asses the distribution of age
ggplot(ab1, aes(x=age)) + geom_bar(color = "black", fill = "red")

```
As we see from the bar chart below, the age of the abalones models a left skewed normal distribution curve with the average/ most common ages centering around the 9-12 age range. 


### Question 2

```{r, include = TRUE}
#Setting the seed so that I can reproduce results 
set.seed(113069)

#Splitting the data into a training and testing set using stratified sampling. This is done using a 80/20 split. 
ab_split <- initial_split(ab1, prop = 0.80,
                                strata = age)
ab_train <- training(ab_split)
ab_test <- testing(ab_split)
```

### Question 3

```{r, include = TRUE}
#Using the training data, I have created a recipe to predict the outcome variable age with all other predictor variables except for rings. 

ab_recipe <-recipe(age ~ type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = ab_train) %>% 
              #Step_dummy is used so that the type variable (nominal) becomes usable
              step_dummy(all_nominal_predictors()) %>% 
              #Created the following interactions between different predictors
              step_interact(terms = ~ longest_shell: diameter) %>% 
              step_interact(terms = ~ shucked_weight:shell_weight)%>% 
              #Broke up the type predictor into two separate predictors and then used that to create 
              #interactions with shucked weight
              step_interact(terms = ~ type_I:shucked_weight) %>%
              step_interact(terms = ~ type_M:shucked_weight) %>%
              #Predictors are centered and scaled
              step_center(all_predictors()) %>%
              step_scale(all_predictors())

summary(ab_recipe)

```
We should not use rings to predict age because age is dependent/ calculated using rings. 


### Question 4

```{r, include=TRUE}
#Creates and stored a linear regression object called lm_model using the lm engine 
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### Question 5

```{r, include=TRUE}
#Set up an empty workflow
lm_wflow <- workflow() %>%
  #Add lm_model created previously 
  add_model(lm_model) %>% 
  #Add recipe created previously 
  add_recipe(ab_recipe)

```

### Question 6

```{r, include=TRUE}
#Create a model to help predict age of abalones 
lm_fit <- fit(lm_wflow, data = ab_train)

#Create a new data frame with the desired values for out prediction. 
new <- data.frame(type = "I", longest_shell = 0.5, diameter = 0.1, height = 0.3, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)

#Predict the age of the abalone based on the above values 
abalone_pred <- predict(lm_fit,new_data = new)
abalone_pred

```


### Question 7

```{r, include=TRUE}
#Use predict to create a tibble of our predicted values 
abalone_train_res <- predict(lm_fit, new_data = ab_train %>% select(-age))
abalone_train_res %>% 
  head()

```
```{r, include=TRUE}

#Use bing_cols below to add the actual observed ages alongside our predicted values
abalone_train_res <- bind_cols(abalone_train_res, ab_train %>% select(age))
abalone_train_res %>% 
  head()

```

```{r, include = TRUE}
#Create a metric that includes R2, RMSE and MAE
ab_metrics <- metric_set(rmse, rsq, mae)
#Apply the metric set to the tibble 
ab_metrics(abalone_train_res, truth = age, estimate = .pred)
```

The metric above signifies a R-squared value of 0.5548639. This indicates that around 55% of the variability observed in the age of abalone is explained by the regression model. This therefore tells us that the fitted model I have coded above does not fit the data extremely well. Thus our model may not predict with the highest level of accuracy any new observations. 
